SBAT: A Simple Browser-based Annotation Tool with Interactive Visualization and Github Support

### 2024-09-05 paper/main.tex
update main part

Based on Brat, the annotation UI of SBAT is responsive, interactive and intuitive to use. To add an annotation, one can simply double-click or select a span with his mouse, and then select corresponding label in the pop-up window. All added labels will be visualized with tags above the annotated text, and different tags differentiate with each other by the user-defined color. And if the user has added comments, they will be shown with a mouse hover. The size and display of the interface can be adjusted by the option button, and the labels and their style are defined in the JSON config file. Besides the main component for annotating text, there is also a step-by-step guide in the beginning for github authentication as well as the export function, all with clear instructions.

@@

SBAT is browser-based, which means one can simply open the website and start annotating in the browser. But unlike Brat, all logics of SBAT are completely executed in the browser, there is no need of software installation or server configuration. One can either store its front-end files in a Github repository to access it via Github pages online, or download the those files to start it locally in the browser. Due to its simple structure, it is fast to run and easy to adapt for different purposes.

@@

SBAT can not only be hosted on Github Pages, but also supports import files from or export annotations to Github repositories. To do so, the user only has to enter his Github Personal Access Token (information about the repository name as well as the its owner's user name are entered in advance under "admin config" in the JSON config file), and can then select the github branch as well as the file to open. For importing file, all file formats are supported, and they will be treated as txt file, i.e. the whole content will be displayed as annotation-ready text in the interface. The only exception is JSON file which has been previously exported from SBAT (which has attributes like text, tokens, entities...) -- Their content will be parsed and the annotations saved in the file will be visualized. To export annotations, the user can optionally leave a commit message and click the "commit" button, a JSON file with original texts and annotations will be commited to the same Github repository from which the file was originally imported. By making use of Github commit, the versions of annotations can be well documented and managed.

@@

SBAT's support for Github also provides possibility for user management in annotation projects. For instance, the admin can create a Github repository for an annotation project and add the repository and owner information in SBAT's config file. By sharing access to this repository with different users, they can contribute to the same project at the same time. Moreover, different branches can be created for different users to create and compare multiple versions. This utilization of Github's file management system facilitates the colloboration of multiple annotators in an annotation project.

@@

\subsection{Annotation Interface}
SBAT's annotation interface is rooted in the frontend component from Brat, which is implemented using XHTML, Scalable Vector Graphics (SVG), JavaScript, and jQuery. On top of that, to remove the dependance on a server, all user actions such as creating, editing or deleting annotations are handled by the browser. This is realized by the LocalAjax class, adapting the LocalAjax function from brat-frontend-editor. It locally simulates AJAX requests to process the data without actually sending real network requests. Other UI elements, such as the login buttons, are also created by HTML and Asynchronous JavaScript. In this way, the complete user interface is embedded in a HTML webpage and all functions are locally executed in the browser.

@@

As SBAT is a static web app without server communications, it can be hosted on Github Pages for free. But it is not mandatory to host the app in order to use it, as the index.js file can be used as an entry point to start the app locally in the browser. If the users want to access it online instead of locally, the source code can be directly uploaded to a Github repository to host a copy of SBAT on Github Pages. Moreover, to cater to different use cases, it is also easy to modify any part of the app -- SBAT uses ECMAScript Modules instead of CommonJS, so all modules are natively loaded in browser and no module bundler such as webpack is used. This allows direct modification of individual modules without the need of re-bundling.

@@

SBAT uses Octokit \cite{octokit.js}, the official SDKs for the Github API, which enables the programmatical implementation of Github operations. The user authentication through Github Personal Access Token enables creating a new user-authenticated Octokit object and obtaining information about branches and file content from the configured repository. Likewise, creating a new commit and pushing to github repository is realized by first obtaining tree of all files from last commit, modifying the changed file or creating a new file, and the creating the new tree and commit it with the REST API endpoint of octokit.

\section{Conclusion}
SBAT is an annotation tool that can be used to attach labels to linguistic data. It is completely browser-based and only requires a browser to be executed. Furthermore, it can be combined well with github to import from and export annotations to user's own github repositories. It aims to provide a simpler and no-server solution for annotation projects.\\

>>>>

Built upon Brat, the annotation editor of SBAT is responsive, interactive and intuitive. To add an annotation, one can simply double-click or select a span with his mouse, and then select the corresponding label in the pop-up menu. In the editor, all annotated spans are highlighted with different color, and the added labels are visualized in tags above the annotated span. Other user-added comments will become visible with a mouse hover. Moreover, the labels and their style can be defined by users in the JSON config file, and the size and display of the editor can be adjusted by the option button. Besides the main annotation editor, there is also an area for Github authentication as well as for the the annotation exportation, all guided with clear instructions.

@@

SBAT is browser-based, which means one can access its website and then simply start annotating in the browser. But unlike Brat, all logics of SBAT are executed in the browser, so there is no need for software installation or server configuration. Due to its simple construction, it is fast to start and convenient to adapt for different purposes.

@@

SBAT enables the users to import files from or export annotations to the Github repositories that they have access to. To do so, one only has to enter his Github Personal Access Token, and can then select the branch and the filename to open the file in the annotation editor. Information about the repository name as well as the repository owner's username should be entered in advance under "admin config" in the JSON config file. To export annotations, the user can optionally enter a commit message in the input field and then click the "commit" button. By making use of Github commit, the versions of annotations can be well documented and managed.\\
For importing files, all file formats are supported, and they will be treated as plain text files, i.e. the complete content will be displayed as plain text in the editor for annotation. The only exception is the JSON file which has been previously exported from SBAT (which has attributes like text, tokens, entities...) -- Their content will be parsed, and the annotations which have been previously saved in the file will be visualized in the editor. And after the user hits the commit button, a JSON file with the original text and its annotations will be committed to the same Github repository from which the file was originally imported.

@@

SBAT's support for Github also enables user management within annotation projects. To accept contributions from multiple annotators in an annotation project, the organizer can create a Github repository for the project, and add the repository and owner information in SBAT's config file. Then, by sharing access to this repository with members of the annotation project, they can all contribute to the same project. Moreover, different branches can be created for different users to create and compare multiple versions. The utilization of Github's collaboration features facilitates management of annotation projects.

@@

\subsection{Annotation UI}
SBAT's annotation editor is rooted in the frontend component from Brat, which is implemented using XHTML, Scalable Vector Graphics (SVG), JavaScript, and jQuery. On top of that, to remove Brat's server-side dependence, the user operations in the editor, such as creating, editing or deleting annotations, are all handled by the browser. This is realized by the LocalAjax class, adapting the LocalAjax function from brat-frontend-editor, which locally simulates AJAX requests to process the data without actually sending real network requests. Other UI functions, such as the user authentication, are also implemented natively in JavaScript. In this way, all components of the user interface are embedded in a HTML webpage and their functionalities are locally executed in the browser.

@@

As SBAT is a static web app without server communications, it can be hosted on Github Pages for free. It is not mandatory, though, to host the app on the internet in order to use it, as the index.js file can be used as an entry point to start the app locally in the browser. Nonetheless, if the user plans to access it online instead of locally, he can directly upload the source code to his Github repository to host a copy of SBAT on his Github Pages. The link can then be shared with others for online access. Moreover, to cater to different use cases, it is also easy to modify any part of the app -- SBAT uses ECMAScript Modules instead of CommonJS, so all modules are natively loaded in the browser and no module bundler such as webpack is used. This allows direct modification of individual modules without the need for re-bundling.

@@

SBAT uses Octokit \cite{octokit.js}, the official SDKs for the Github API, which enables the programmatical implementation of Github operations. The user authentication through Github Personal Access Token creates a new user-authenticated Octokit object, which can then be used to obtain information about branches and file content from the user-defined repository. Likewise, committing annotations to a Github repository is realized by first obtaining the tree of all files from the last commit in the repository, among the files modifying the changed file or creating a new file, and finally creating a new tree with the updated files and committing it with the REST API endpoint of Octokit.

\section{Conclusion}
SBAT is an annotation tool for attaching labels to linguistic data, which aims to provide a simpler and no-server solution for annotation projects. It is completely browser-based and only requires a browser for usage. It retains the essential features of Brat's annotation editor, such as annotation operations and visualizations, while offering integration with Github, enabling user authentication and the import and export of annotations from and to Github repositories.\\
